\artus -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-09-09 Core/doc/latex/paper.tex
further work

At the beginning a configuration file in the JSON format is read in. It contains information on the input data and settings for the pipelines and processors described in the following. Each of the modules described in the following has access to the settings read in from the configuration file.

@@

A pipeline consists of an arbitrary number of processor modules that are run in a sequence. Three kinds of processors are distinguished:
\begin{description}
\item[Producers] are meant to calculate new quantities based on the information in the event and the product and following the settings of the pipeline.
\item[Filters] apply cuts. In case an event does not survive a requirement defined by a filter, the processing of the subsequent producers and filters is skipped.
\item[Consumers] can be added to local pipelines. They are executed after the producers and filters. The task of consumers is to take quantities from the event and the product and store them in the output according to the pipeline settings.
\end{description}
Producers and filters can be configured in an arbitrary order. Consumers are executed after the chain of producers and filters. The output together with the complete configuration is written to a ROOT file. These processors are created and managed by a dedicated factory according to the configuration.

This structure extends the established concepts of existing frameworks like FWLite~\cite{FWLite} from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writeable parts.
\item[Re-utilisation of code] Processors can be shared among different analysis groups. Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability] Each module uses settings that are read from the the configuration. Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance] The concept of pipelines allows to analyse the data with different configurations at the same time, while the data is only read in only once and common processing steps are shared in the global pipeline. Therefore the cost-intensive input operations are reduced to a minimum.
\end{description}

\section{Building an End-user Analysis \label{section_artus_analysis}}

In the \artus framework each concrete analysis is defined by a set of object types. It needs an implementation of the classes for settings, the event and the product. The settings class defines the tags that should be able to be read from the configuration file and the corresponding types of the values. The event class defines the event content being read from the input and the product class contains members for new quantities being calculated by the analysis. These three classes need to be derived from the base classes in the core of the framework.

@@

The most important part of an analysis are the processors. The abstract processors of the \artus core need to be implemented to realise the physics needs of the analysis in the form of code. New processors have to be registered in the factory. After the basic structure of an analysis is set up, most of changes and additions are only related to the processors.

>>>>

\artus respects the separation of code and settings. Settings select what has to be done and with wich parameters. The code does perform these steps, ideally without the user being fully aware of implementation details. The single source of truth for an analysis is therefore a JSON file that completely defines the analysis. It contains information on the input data and settings for the pipelines and processors described in the following. Each of the modules described in the following has access to the settings read in from the configuration file.
The \artus package comes with sophisticated tools written in python, that help in generating this JSON file.

@@

A pipeline consists of an arbitrary number of processor modules that are run in a sequence. Three kinds of processors are distinguished:
\begin{description}
\item[Producers] are meant to calculate new quantities based on the information in the event and the product and following the settings of the pipeline.
\item[Filters] apply a selection based on event properties. In case an event does not survive a requirement defined by a filter, the processing of the subsequent producers and filters can be skipped.
\item[Consumers] can be added to local pipelines. They are executed after the producers and filters. The task of consumers is to take quantities from the event and the product and store them in the output according to the pipeline settings.
\end{description}
Producers and filters can be configured in an arbitrary order. Consumers are executed after the chain of producers and filters. The output together with the complete configuration is written to a ROOT \todo{Referenz auf ROOT?} file. These processors are created and managed by a dedicated factory according to the configuration.
\todo{I would not put the remaining part of this chapter here. It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced}
This structure extends the established concepts of existing frameworks like FWLite~\cite{FWLite} from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writeable parts.
\item[Re-utilisation of code] Processors can be shared among different analysis groups. Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability] Each module uses settings that are read from the the configuration. Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance] The concept of pipelines allows to analyse the data with different configurations at the same time, while the data is only read in only once and common processing steps are shared in the global pipeline. Therefore the cost-intensive input operations are reduced to a minimum.
\end{description}

\section{Building an End-user Analysis \label{section_artus_analysis}}
\todo{If I wouldn't know the structure already, the following part wouldn't have helped my. Either we support it with a drawing and more detailled explanations or we just refer to a readme...}
In the \artus framework each concrete analysis is defined by a set of object types. It needs an implementation of the classes for settings, the event and the product. The settings class defines the tags that should be able to be read from the configuration file and the corresponding types of the values. The event class defines the event content being read from the input and the product class contains members for new quantities being calculated by the analysis. These three classes need to be derived from the base classes in the core of the framework.

@@

The most important part of an analysis are the processors. The abstract processors of the \artus core need to be implemented to realise the physics needs of the analysis in the form of code. New processors have to be registered in the factory. After the basic structure of an analysis is set up, most changes and additions are only related to the processors.
